\textbf{\Large Development and comparison of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

### 2024-02-19 manuscript_BiSC_outdoor_AP.tex
finished paper for co-authors

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Title of the paper --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}

@@

%%% --- Authors --- %%%
\author[1, 2, 3]{Alan Domínguez}
\author[1, 3, 4]{Payam Dadvand}
\author[1]{Marta Cirach}
\author[1]{Bruno Raimbault}
\author[1]{Toni Galmes}
\author[1]{Karl Samuelsson}
\author[1, 2, 3]{Mark Nieuwenhuijsen}
\author[1, 2, 3]{Jordi Sunyer}
\author[1, 2, 3]{Xavier Basagaña}
\author[1, 2, 3]{Ioar Rivas}

%%% --- Affiliations --- %%%
\affil[1]{Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
\affil[2]{Universitat Pompeu Fabra (UPF), Barcelona, Spain.}
\affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
\affil[4]{London School of Hygiene and Tropical Medicine (LSHTM), London, UK.}

@@

Air pollution is the leading environmental risk factor for human health worldwide linked to a wide range of adverse health outcomes. A detailed and comprehensive assessment of outdoor air pollution exposure considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility during pregnancy to traffic-related air pollution. Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses during pregnancy for several traffic-related air pollutants; NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn). The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona. Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO2 between 2018-2021. Additionally, a second campaign was conducted, named "BiSCAPE", and included four distinct measurement periods between 2021-2022. LUR and DM were constructed using conventional state-of-the-art techniques, while HM was designed by harnessing Random Forest algorithm, integrating outputs from DM with explanatory variables from the LUR model, monitoring stations and detailed traffic counts. Model performance was evaluated using a cross-validation procedure for LUR and HM, and by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations. The models demonstrated moderate to good performance for LUR, DM and HM. HM outperformed LUR and DM models for NO\textsubscript{2} (R\textsuperscript{2} = 0.64), PM\textsubscript{2.5} (R\textsuperscript{2} = 0.66), and BC (R\textsuperscript{2} = 0.86). LUR models show better performance for PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = 0.89), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = 0.87), PM\textsubscript{2.5 - Zn} (R\textsubscript{2} = 0.85). Correlations between the air pollutant estimates for the different models were moderate to strong for both long-and-short term exposures. Long-term exposure (entire pregnancy) at the residential addresses exhibited correlations ranging from rs = 0.48 to 0.92, while for short-term exposure (weekly exposure) varied from rs = 0.37 to 0.90. Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates during pregnancy. The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy.

@@

While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains (multi-country and national domain models), these models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}. This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to capture the differences in exposure for the population living in relatively small areas (i.e. city compared to country-level) to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Dispersion model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Dispersion model evaluation}
Briefly, once the input data has been introduced to the ADMS Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out. Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others. Additionally, we compared the results of the model with data from the 4 campaigns carried out at BiSCAPE sites. In total, 98 periods were used to evaluate the performance of NO\textsubscript{2} model, 66 periods for PM\textsubscript{2.5}, and 93 periods for BC. The closer these values are, the better the model's performance. Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Hybrid model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hybrid model evaluation}
We assessed the performance of each hybrid model by contrasting predictions with observations. Two distinct validation approaches were employed. Firstly, \textit{10-fold cross-validation (10-fold CV} was used to provide an internal validation measure, where the model was trained on 80\% of the data. Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance against previously unseen data. While both methods differ in their approach to partitioning the data, the fundamental distinction lies in how the data is utilized during the training and validation phases. The \textit{10-fold CV} method systematically rotates through the data, ensuring that each subset serves as a validation set once, while the external validation approach provides a straightforward assessment against a separate data portion. Additionally, we employed a grid search for hyperparameter tuning to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Comparison between model estimates --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison between model estimates}
Exposure estimates obtained from LUR, DM, and HM were compared at the BiSC participant's address level for long- and short-term exposure. We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman (Rho) correlation coefficients. We Bland-Altmant plots to investigate the level of agreement between the different types of models.

@@

\Cref{fig6} shown Bland-Altmant plots to assess the level of agreement over the concentrations ranges between the models.

%%% --- Figure. Comparison/agreement between moodel predictions estimates --- %%%
% Put the figure text closer to the Figure 3

% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!h]
\includegraphics[width=0.7\textwidth]{figures/final_bland_altmant.png}
\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn). using LUR, DM, and HM}}
\label{fig6}
\end{figure}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed outdoor air pollutants models using three approaches (LUR, DM, and HM). Models show moderate to good performance 
evaluated by previously defined metrics and displayed in more detail in the \textbf{supplementary materials}. As expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed. We believe that by combining multiple data sources and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.

\subsection{Comparison between LUR, DM, and Hybrid model}

\subsection{Comparison with Previous studies}

\subsection{Strengths and Limitations}
One of the main strengths of this study is the extensive exposure assessment

@@

Air pollution exposure concentration estimates from LUR, dispersion and hybrid models were moderately to strongly correlated. We observed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Title of the paper --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{\Large Development and comparison of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}}

@@

%%% --- Authors --- %%%
\author[1, 2, 3]{\small Alan Domínguez}
\author[1, 2, 3]{Payam Dadvand}
\author[1, 2, 3]{Marta Cirach}
\author[4]{Gustavo Arevalo}
\author[4]{Lluís Barril}
\author[1, 2, 3, 5]{María Foraster}
\author[1, 2, 3]{Mireia Gascon}
\author[1]{Bruno Raimbault}
\author[1]{Toni Galmes}
\author[1, 2, 3]{Laura Goméz-Herrara}
\author[1, 2, 3]{Cecilia Persavento}
\author[1, 6]{Karl Samuelsson}
\author[4]{Jose Lao}
\author[7]{Teresa Moreno}
\author[7]{Xavier Querol}
\author[1, 2, 3]{Cathryn Tonne}
\author[1, 2, 3]{Mark Nieuwenhuijsen}
\author[1, 2, 3]{Jordi Sunyer}
\author[1, 2, 3, *]{Xavier Basagaña}
\author[1, 2, 3, *]{Ioar Rivas}

%%% --- Affiliations --- %%%
\affil[1]{\small Barcelona Institute for Global Health (ISGlobal), Barcelona, Spain.}
\affil[2]{Universitat Pompeu Fabra (UPF), Barcelona, Spain.}
\affil[3]{CIBER Epidemiología y Salud Pública (CIBERESP), Madrid, Spain.}
\affil[4]{Barcelona Regional}
\affil[5]{Blanquerna School of Health Science, Universitat Ramon Llull (URL), Barcelona, Spain.}
\affil[6]{Department of Building Engineering, Energy Systems and Sustainability Science, Faculty of Engineering and Sustainability Science, University of Gävle, Gävle, Sweden}
\affil[7]{Institute of Environmental Assessment and Water Research (IDAEA-CSIC), Barcelona, Spain.}
\affil[*]{Join last authors}

% --- removing date from manuscript --- % 
\date{}

@@

Air pollution is the leading environmental risk factor for human health worldwide linked to a wide range of adverse health outcomes. A detailed and comprehensive assessment of outdoor air pollution exposure considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility
during developmental periods over the lifespan to traffic-related air pollution (TRAP). Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses for several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn). The overall aim is to obtain accurate weekly estimates of traffic-related air pollution, intended to be used in the exposure assessment for a cohort of pregnant women in the city of Barcelona. Two separate monitoring campaigns were conducted: the "BiSC-home" campaign took place at the home addresses of BiSC participants and involved measuring only NO\textsubscript{2} between 2018-2021. Additionally, a second campaign was conducted, named "BiSCAPE", and included four distinct measurement periods between 2021-2022. LUR and DM were constructed using conventional state-of-the-art techniques, while HM was designed by harnessing Random Forest algorithm, integrating outputs from DM with explanatory variables from the LUR model, monitoring stations and detailed traffic counts. Model performance was evaluated using a cross-validation procedure for LUR and HM, and by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality stations. The models demonstrated moderate to good performance for LUR, DM and HM. HM outperformed LUR and DM models for NO\textsubscript{2} (R\textsuperscript{2} = 0.64), PM\textsubscript{2.5} (R\textsuperscript{2} = 0.66), and BC (R\textsuperscript{2} = 0.86). LUR models show better performance for PM\textsubscript{2.5 - Fe} (R\textsuperscript{2} = 0.89), PM\textsubscript{2.5 - Cu} (R\textsuperscript{2} = 0.87), PM\textsubscript{2.5 - Zn} (R\textsubscript{2} = 0.85). Correlations between the air pollutant estimates for the different models were moderate to strong for both long-and-short term exposures. Long-term exposure (entire pregnancy) at the residential addresses exhibited correlations ranging from rs = 0.48 to 0.92, while for short-term exposure (weekly exposure) varied from rs = 0.37 to 0.90. Our work emphasizes the importance of employing diverse exposure assessment methods to evaluate the exposure to traffic-related air pollutants and to disentangle differences in effect estimates during pregnancy. The exposure estimates from these different models can be used in the BiSC cohort to potentially identify critical windows of susceptibility to air pollution during pregnancy. \\

@@

While significant advancements have been made in spatiotemporal modeling of air pollution over large-scale domains (multi-country and national domain models), these models often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}. This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to capture the differences in exposure for the population living in relatively small areas (i.e. city compared to country-level) to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Dispersion model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\subsection{Dispersion model evaluation}
Briefly, once the input data has been introduced to the ADMS-Urban model, the calculation of NO\textsubscript{2}, PM\textsubscript{2.5}, and BC concentration levels has been carried out. Dispersion models were evaluated by comparing the yielded results from the ADM-Urban model with the real values measured by the air quality monitoring stations (XVPCA), including Ciutadella, Eixample, Gracia, Palau Reial, Poblenou, Sants, among others. Additionally, we compared the results of the model with data from the 4 campaigns carried out at BiSCAPE sites. In total, 98 periods were used to evaluate the performance of NO\textsubscript{2} model, 66 periods for PM\textsubscript{2.5}, and 93 periods for BC. The closer these values are, the better the model's performance. Details from the metrics used for dispersion model validation and the XVPCA stations locations can be found in the \textbf{supplementary material}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Hybrid model evaluation --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hybrid model evaluation}
We assessed the performance of each hybrid model by contrasting predictions with observations. Two distinct validation approaches were employed. Firstly, \textit{10-fold cross-validation (10-fold CV} was used to provide an internal validation measure, where the model was trained on 80\% of the data. Secondly, external validation was performed by allocating 20\% of the data as a test set to further evaluate the model's performance against previously unseen data. While both methods differ in their approach to partitioning the data, the fundamental distinction lies in how the data is utilized during the training and validation phases. The \textit{10-fold CV} method systematically rotates through the data, ensuring that each subset serves as a validation set once, while the external validation approach provides a straightforward assessment against a separate data portion. Additionally, we employed a grid search for hyperparameter tuning to select the best configuration to obtain the optimal model based on the \textit{Root Mean Squared Error (R\textsuperscript{2})}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Comparison between model estimates --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison between model estimates}
Exposure estimates obtained from LUR, DM, and HM were compared at the BiSC participant's address level for long- and short-term exposure. We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman correlation coefficients. Bland-Altmant plots were used to investigate the level of agreement between the different types of models. \\

@@

\textbf{\Cref{fig6}} shown Bland-Altmant plots to assess the level of agreement across the concentrations ranges between the models. The limits of agreement, represented by the red dashed lines  contains the majority of the points with some outliers present in the plots indicating good agreement in the estimated concentration, but with potential disagreement at higher estimated concentrations, showing an upward trend when comparing the model`s estimates. Overall, when comparing outdoor air pollution estimates at address level the models show good agreement at lower and medium ranges, with low bias that is observed as the blue ticked line is close to 0 for almost all comparison between models.

%%% --- Figure. Comparison/agreement between moodel predictions estimates --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!h]
\includegraphics[width=0.7\textwidth]{figures/final_bland_altmant.png}
\caption{Bland-Altmant plots for comparing outdoor air pollution estimates for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, and Zn). using LUR, DM, and HM}}
\label{fig6}
\end{figure}
\vspace*{-\baselineskip}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}

% --- General --- % 
We successfully developed outdoor air pollution models (NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents) to assess 
exposure within the city of Barcelona for the period 2018-2021. Utilizing three different approaches: land use regression, dispersion modeling, and hybrid models were tailored to provide exposure estimates across various time windows releveant to our study population in the BiSC cohort, including entire pregnancy duration, specific trimesters, and weeks of pregnancy for each participant. The performance of these models was moderate to good, as determined by different evaluation metrics. The LUR models accounted for approximately 45\% to 89\% of the variance in pollutant levels, indicating their strength in capturing spatial variability in air pollution concentrations. HM models, which integrate a broader range of data inputs and modeling techniques, explained 44\% to 86\% of the variance. In contrast, DM models, focusing on physical and chemical processes affecting pollutants, explained a lower range of variability, between 26\% and 65\%. These findings align with the outcomes of similar air pollution modeling studies conducted across Europe and specifically within Barcelona \cite{fung2024, mila2023, criado2023, rovira2022, chen2020, dehoogh2018, dehoogh2013}. For both LUR and HM approaches, the selection of predictor variables demonstrated considerable variability across different pollutants, reflecting the influence of distinct emission sources. Notably, BC achieved the highest model performance across all methods, followed by NO\textsubscript{2}, PM\textsubscript{2.5}, and their constituents. This pattern highlights the critical role of source-specific factors in influencing the accuracy and reliability of pollution exposure estimates.\\

% --- NO2 models --- %
The performance of our LUR NO\textsubscript{2} model was higher than previous European models \cite{rivera2012} as well as earlier models specific to Barcelona \cite{criado2023}, indicating an enhanced capacity to capture the intra-urban variability of NO\textsubscript{2} concentrations levels. Notably, our LUR model outperformed a recently developed microscale LUR model, which was part of a data fusion method designed to minimize prediction uncertainty by integrating a LUR base map with hourly dispersion estimates from the CALLIOPE-Urban model within Barcelona \cite{criado2023}. The predictor variables of our model were predominantly traffic-related, and were selected due to their strong association with NO\textsubscript{2} emissions \cite{casallas2018}. Conversely, our DM model exhibits a lower performance compared to the CALLIOPE-Urban system, a regional-to-street-scale modelling framework that offers high-spatial-and-high-temporal-resolution (10m x 10m, hourly) by integrating the mesoscale airquality forecasting system CALIOPE (WRF-HERMES-CMAQ-BSS-DREAM8b) with an urban roadwide dispersion model (R-LINE) \cite{benavides2019}. Regarding our HM, where we integrated detailed information on traffic counts across the city of Barcelona, air pollution and meteorological data from monitoring stations, and estimates from the DM model through a RF algorithm, we achieved better performance compared to both our LUR and DM models. Most relevant predictor variables were estimates from the DM, monitoring station, traffic counts, solar radiation traffic-related spatial variables. We attribute the enhanced variance explanation by the HM to two main factors. Firstly, the incorporation of outputs from the DM model allows to improve the spatiotemporal variability in our model \cite{michanowicz2016, korek2017}, in addition traffic information allows to capture other emission sources not previously captured \cite{pinto2020} in our other two approaches. Secondly, the integration of the RF algorithm enables the model to account for non-linear relationships and potential interactions between predictor variables that the LUR and DM models did not fully exploit \cite{araki2018, chen2019}. Despite these advancements, a recent hybrid model developed for the entire region of Catalonia, which utilized information on air pollution measurements from monitoring stations, land use variables, and satellite remote sensing products, outperformed the performance of our HM for NO\textsubscript{2} \cite{mila2023}. This improvement is likely attributable to the use of data from the TROPOspheric Monitoring Instrument (TROPOMI), providing deeper insights into the vertical atmospheric distribution of NO\textsubscript{2} \cite{petetin2023}.\\

% --- PM25 particles --- % add predictor variables 
PM\textsubscript{2.5} have a moderate performance in LUR models similar to  previous models that aim to predict within-city variability in spanish cities \cite{rivera2012, wang2014} but lower than other multi-cities LUR models developed in Europe \cite{eeftens2012, gulliver2018}. Predictor variables used in the LUR model were mainly traffic, road density, and residential variables. We believe that the performance obtained from our LUR model can only explain a part of the variation of the  PM\textsubscript{2.5} concentrations across Barcelona. One potential cause that could partially explain the moderate performance of our LUR model is that PM\textsubscript{2.5} in Barcelona came from nine to ten major contribution sources that have been previously identified in two chemical species and source contribution studies for PM\textsubscript{2.5} \cite{in2021, amato2024}, we believe that our model was not able to take into account the variability of this sources since LUR models mainly use traffic-related variables to predict PM\textsubscript{2.5} concentrations levels, and previously have been described that LUR models are less effective to predict sources other than traffic \cite{dehoogh2014}. The performance of our DM model was less effective than the one obtained from the LUR model. One potential explanation could be the limited coarser spatial resolution from DM compared to LUR. Our lUR model uses observed data and land use characteristics to estimate air pollution levels, potentially offering a finer spatial resolution that better captures local variations, especially within urban settings \cite{hoek2017}. Since there were no previously available published DM models PM\textsubscript{2.5} developed in the city of Barcelona or at a Catalunya scale we were not able to compare the performance. The enhanced performance of our PM\textsubscript{2.5} HM model compared to our previous approaches can largely be attributed to the incorporation of comprehensive information during the model-building process and  the employment of the RF algorithm significantly increased our model's accuracy in predicting air pollution levels. This approach enabled us to account for additional sources of variability in PM\textsubscript{2.5} concentration levels, identifying emissions sources not captured by the other models. Additionally, our HM model demonstrated improved performance metrics compared to a previously developed RF model \cite{mila2023}. Regardless of these advancements in modeling PM\textsubscript{2.5} concentrations by our HM, it was found that recent European multi-stage HM models, utilizing RF algorithm for predicting daily PM\textsubscript{2.5} levels at a country-level with a high spatiotemporal resolution, exhibit superior performance metrics. These models integrated satellite Aerosol Optical Depth (AOD) measurements along with PM and meteorological data monitored at various locations and spatial predictors at the grid cell level, enhancing predictive accuracy beyond our HM model \cite{dehoogh2018, stafoggia2019, stafoggia2020, schneider2020}. This difference in model performance can be attributed to two main factors. First, the comprehensive exposure assessment period employed by these models, which utilized routine several years of data from routine monitoring stations to more effectively capture temporal trends in air pollution exposure levels. Second, the inclusion of AOD measurements provided additional insights into the background and regional variability of air pollution levels with higher correlations in Western Europe compared to the rest of the Region \cite{chen2024}.\\

% --- Black Carbon (BC) --- %
BC models have the highest performance across all pollutants, this was systematically observed in all our exposure assessment methods explored. 
The LUR-developed model has greater performance than previous BC models developed in Europe \cite{hoek2015, van2020}. However, there are only a few studies that aim to predict BC concentrations levels across Europe, this limited our comparison with other models. Our HM outperformed a previous hybrid European-wide model developed using large-scale satellite data, chemical transport models, and small-scale and land use predictors to predict BC \cite{dehoogh2018}, and a hybrid LUR model that 
to estimate BC and metal components across the metropolitan area of Pittsburgh, USA \cite{tripathy2019}. Predictor variables used for the development of LUR and HM models were mainly traffic, residential, and population density-related. We believe that the high performance in both LUR and HM is that BC levels proportionally reflect traffic-related process  \\

% --- PM2.5 constituents particles --- % 
LUR has a very good performance in predicting PM\textsubscript{2.5} constituents pollutants, while HM has a moderate performance. In both LUR and HM models, traffic-related variables emerged as significant predictors for PM\textsubscript{2.5} constituents, this was consistent with results obtained from stepwise and RF analysis. This aligns with previous evidence indicating the influence of traffic-related variables that explain a large proportion of the variability in the measured concentrations \cite{dehoogh2013, chen2020}. Additionally, industry-related variables appears as relevant predictors in predicting PM\textsubscript{2.5 - Cu} and \textsubscript{2.5 - Zn} in both LUR and HM, possibly reflecting emissions from industry sector \cite{belis2013, chen2020}. Our PM\textsubscript{2.5} constituents models outperformed in performance to those published in previous studies \cite{}. Differences in performance between LUR and HM models in predicting PM\textsubscript{2.5} constituents could be explained due to stability for long-term concentrations in PM\textsubscript{2.5} composition.\\

% --- Predicted concentrations ranges --- %

% --- Correlations and agreement between models --- %
Comparisons of air pollution estimates at the address level indicated that the correlations between the different models' estimates were moderate to strong for both long-term and short-term exposure periods. For NO\textsubscript{2}, PM\textsubscript{2.5} and BC we observed a positive correlation for both long-and-short exposures. Long-term correlations were strong for NO\textsubscript{2}, PM\textsubscript{2.5} and BC in comparison to short-term correlations. This is mainly due to a smoothing effect over the pregnancy period that could lead to more stable correlations between pollutants and also a seasonality effect between trimesters \cite{wang2018, morgan2023, hao2023}. This was not observed for PM\textsubscript{2.5} constituents, on the contrary, we observed the reverse pattern in correlation showing stronger correlations for short-term exposure (weekly exposures). Correlation between pollutants estimated for the same exposure assessment method was higher for DM, this pattern was also observed in  a previous study \cite{klompmaker2021}. One potential explanation could be that DM models often rely on the same emission source inventory to predict different air pollutants that could lead to higher correlations between different pollutants in comparison to other approaches such as LUR and HM models. PM\textsubscript{2.5} constituents were positive correlated between them for both long- and-short term windows of exposure for LUR and HM models. PM\textsubscript{2.5 - Zn} estimated using LUR and HM models were poorly correlated against estimated pollutants (NO\textsubscript{2}, PM\textsubscript{2.5}, and BC) from other exposure assessment methods, this pattern was observed for long-term and short-term exposure but was more pronounced for HM estimated pollutants. We believe that this results could be partially explained since both PM\textsubscript{2.5 - Fe} and PM\textsubscript{2.5 - Cu} come mainly from traffic, particularly from venhicle non-exhaust factors, while PM\textsubscript{2.5 - Zn} could came from industry emission sources as well this could explain the magnitude of some correlations in \textbf{Fig 5.A} and \textbf{Fig 5.B} \cite{amato2024}. Overall, there was good agreement among the LUR, DM, and HM model estimates for all pollutants. However, the lowest agreement was observed for higher air pollution estimates, suggesting that at elevated pollution levels, discrepancies between model estimates become more pronounced. Our reported findings were consistent with similar studies, which report similar performance metrics for the outdoor air pollutants assessed and relatively good agreement across different \cite{mila2023} and models exposure assessment methods \\

% --- Strengths --- %
Our study has several strengths. First, we use three different approaches to develop outdoor air pollution models for different TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn), allowing for a comprehensive assessment of the effects of air pollution in further epidemiological studies. Secondly, we take into account the potential measurement error introduced by the COVID pandemic in the estimations of our models by incorporating detailed information from pre-and- post pandemic traffic across the city, XVPCA monitoring station network, and traffic scenarios, allowing us to capture changes in air pollution trends across Barcelona. Thirdly, we report different metrics to assess the performance of outdoor air pollution models that are commonly employed to assess performance in different approaches to estimate air pollution levels. Additionally, we report the results for DM models which are frequently overlooked in the exposure assessment components of epidemiological studies. Fourthly, our LUR and HM models had the advantage of a large training dataset in measured concentrations by combining measurements from the BiSC-home and BiSCAPE monitoring campaign with different contrasts in exposure levels across Barcelona, enhancing the robustness and depth of our analysis. Finally, to our knowledge, this is one of the first studies to comprehensively assess different TRAP's using a multifaceted modeling strategy that spans both pre- and post-pandemic periods. It assesses their relative performance at an urban scale, enabling the disentanglement of differences in effect estimates derived from distinct exposure assessment methodologies. This approach allows for a detailed examination of within-city variability in exposure, contributing valuable insights into further epidemiological studies.\\

% --- Limitations --- % 
The results obtained in this study, which stands as one of the first comprehensive assessments of TRAPs, must be interpreted within the context of its limitations. 
First, the methodology employed in the construction of our different models difficult direct comparisons between models, thus making it challenging to draw strong conclusions about their relative performance and accuracy. This inherent complexity in model comparison underscores the need for cautious interpretation of our results. Second, the DM was designed to predict air pollution levels at specific locations and within particular temporal frames. This specificity limits the reproducibility and broader applicability of the HM, which rely on dispersion estimates as a fundamental component of their predictive variables. However, this does not limit the potential utility of our LUR models that could be extended and used in  following epidemiological studies, out of the initial period for what the model was built. These models, despite the study's limitations, offer valuable insights and could be adapted for use in future epidemiological research beyond the initial scope for which they were developed. Third, the absence of external validation data restricts our capacity to evaluate  the accuracy of our preferred modelling approach. Despite this, we assess the performance of the NO\textsubscript{2} model given some lights regarding the most suitable approach for our context. However, we can no extending these findings to other pollutants due to differences in their sources, behavior, and environmental interactions. Despite these challenges, our study lays the groundwork for future research in exposure assessment in urban areas, providing a basis for methodological refinement and encouraging the development of more generalized models that can be  externally validated.

@@

In conclusion, our study innovatively developed and compared three different exposure assessment methods to estimate outdoor air pollution exposures to several TRAPs: NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn) within the urban context of Barcelona. Our work provides a comprehensive evaluation of outdoor air pollution exposure, highlighting the critical role of diverse exposure assessment methods in understanding the effects and discrepancies of TRAPs exposure. Finally, our exposure estimates derived from the models developed will be used in the BiSC cohort and potentially in further epidemiological studies in Barcelona and its metropolitan area.
